\heada{SOLUTION}{RESTart}
\hspace{1.0cm}{{rest,$<$fileext$>$ \hfill}}
\headb

A restart may be made using the results from previous
analyses (which are retained in the restart read file specified
at the start of each analysis).  After entering the
command language program the restart may be specified.  If the previously
computed problem was "dynamic", it is necessary to
specify the {\tt TRAN}sient command prior to issuing a {\tt REST}art
command in order to restore the velocity and acceleration
states.  If the previous problem was static and the new
analysis is to be continued as a dynamic calculation, the
{\tt REST}art is issued before the {\tt TRAN}sient command (since the
previous analysis did not write a velocity or acceleration
state to the restart file.

The {\tt fileext} option is used to restart with files generated
using the {\tt SAVE} command with the same specified {\tt fileext}.

The use of the restart option requires considerable
care to ensure that the previous results used are proper.
At the termination of any analysis which computes a solution
state a new file is saved on the restart write file specified
at the start of the analysis.  If the last analysis
performed is for a different problem than the current one an
error will result.

If no new solution state is computed during command language execution
(e.g., only plotting is performed) no restart file is
written to the specified fileset - the previous restart file
is retained on the original fileset.
\vfill\eject
