\chapter[Problem Definition]{PROBLEM DEFINITION}
\label{def}

To perform an analysis using the finite element method the first step is
to subdivide the region of interest into elements and nodes.  In this
process the analyst must make a choice on: (a) the type of elements to
use, (b) where to place nodes, (c) how to apply the loading and
boundary restraints, (d) the appropriate material model
and values of its parameters in each element, and (e) any other aspects
relating to the particular problem.  The specification of the node and
element data defines what we will subsequently refer to as the
{\it finite element mesh} or, for short, the {\it mesh} of the problem.  
In order to complete a problem specification it is necessary also to
specify additional data, e.g., boundary conditions, loads, etc..

Once the analyst has defined a model of the problem to be solved it is 
necessary to  define the nodal and element data in a form
which may be interpreted by {\sl FEAP}.  The steps to define 
a mesh for {\sl FEAP}
are contained in Chapters \ref{meshin} to \ref{manip}
and the input data for several
example problems is described in the {\sl Example Manual}.
Each of the commands available for constructing the mesh data is described
in Appendix A.

The second phase of a finite element analysis is to specify the 
solution algorithm for the problem.  This may range from a simple
linear static (steady state)
analysis for one loading condition to a complicated transient
non-linear analysis subjected to a variety of loading conditions.  {\sl FEAP}
permits the user to specify the solution algorithm utilizing
a solution command language which is described in Chapter \ref{command} and
also illustrated in the {\sl Example Manual}.
Each solution command is also described in Appendix B of the report.

\section{Execution of FEAP and Input/Output Files.}

The execution of {\sl FEAP} is initiated by issuing the command:
\vskip 0.1in \par\noindent
\begin{verbatim}
            FEAP
\end{verbatim}
In PC use it is possilble to execute the program using standard windows options
or to open an MS-DOS window and execute with the above command.
If this is a first execution of the program it is necessary to
provide names for the file containing
the input data and those to receive output information.  Upon a successful first
execution of the program a file {\tt feapname} will be written to disk to
preserve the name for each of the input and output disk files.
If it is desired to reinstall the program the {\tt feapname}
file should be deleted and the {\tt FEAP} command then reissued.

For each subsequent execution of the program using the {\tt FEAP} command,
the analyst receives prompts for a new input data filename,
as well as for the filenames which are to contain the
output of results and diagnostics, and
restart files (used if subsequent analyses are desired starting 
with the final results of a previous execution).
The name of a default selection will also be indicated and may be accepted
by pressing the return (enter) key without specifying any other data.
Prior to running {\sl FEAP} it is necessary to create the input data file
using a standard text editor or word processing system. The other
files are created by {\sl FEAP}.  A large part of the remainder of
this manual is directed to defining
the steps needed to create a valid input data file and to describe the
command language instructions needed to solve and output results for
a problem.

Execution of {\sl FEAP} also may be made without specifying filenames 
interactively.  The command line to perform this mode of execution is:
\begin{verbatim}
       feap -iIfile -oOfile -rRfile -sSfile -pPfile
\end{verbatim}
Each parameter defines the name of the file which either contains input
data or will be used to produce the output data.  The files are:
\begin{verbatim}
       i = input            : Ifile is file containing input
                              data
       o = output           : Ofile is file for outputs
       r = restart read file: Rfile is filename
       s = restart save file: Sfile is filename
       p = plot             : Pfile is root name for file
                              containing time history data.
\end{verbatim}
Except for the name of the input data file, these parameters are optional.
Thus, the minimum command line for an execution is:
\begin{verbatim}
       feap -iIfile
\end{verbatim}
the other files will be given by replacing the first character
in the Ifile name by {\tt O}, {\tt R}, {\tt S}, {\tt P}.

\begin{description}
\item
Note: There can be NO blank characters between the {\tt -i}, {\tt -o}, etc.
and the corresponding file name.  That is
\begin{verbatim}
       feap -i Ifile
\end{verbatim}
will cause an error.
\end{description}

\section{Modification of Default Options}

At the time that the executable version of {\sl FEAP} is created
default values for several parameters may be set in file {\tt feap71.f}.
These default parameters may be changed without recompilation
by creating a file named {\tt feap.ins} which contains the new values for
specific parameters.  This file must be placed in each directory where
problems are to be solved.  The {\tt feap.ins} file contains separate records
which define the default parameters to be employed during any solution.  The
current options are given in Table \ref{tab22}.

\begin{table}
\begin{center}
\begin{tabular}{l l l |  l}
Option & Parameter 1 & Parameter 2 & Description \\ \hline
manfile & mesh & path & Path to locate MESH \\
 & & & COMMAND manual pages \\
        & macr & path & Path to locate SOLUTION \\
 & & & COMMAND manual pages \\
        & plot & path & Path to locate PLOT \\
 & & & COMMAND manual pages \\
        & elem & path & Path to locate USER \\
 & & & ELEMENT manual pages \\ \hline
noparse & & & Assumes input data is mostly \\
 & & & numeric  \\
parse & & & Assumes input data contains \\
 & & & parameters  \\ \hline
graphic & prompt & off & Turns off contour prompts \\
        &        & on  & Turns on contour prompts \\
        & default & off & Turns off graphics defaults \\
        &         & on  & Turns on graphics defaults \\ \hline
postscr & color & reverse & Makes color PostScript files \\
 & & & with reversed order. \\
        &       & normal  & Makes color PostScript files \\
 & & & with normal order. \\ \hline
helplev & basic & & Default level for commands \\
 & & & Same as: MANU,0 \\
        & interm & & Default level for commands \\
 & & & Same as: MANU,1 \\
        & advance & & Default level for commands \\
 & & & Same as: MANU,2 \\
        & expert & & Default level for commands \\
 & & & Same as: MANU,3 \\
increment & value & & Set increment value change to \\
 & & & force reduction in array size. \\
\end{tabular}
\caption{Options for Changing Default Parameters}
\label{tab22}
\end{center}
\end{table}
