\chapter[Mesh Input Data]{MESH INPUT DATA SPECIFICATIONS}
\label{meshin}

The description of the mesh data for a problem to be solved by {\sl FEAP}
consists of several parts as described in the following sections.

\section{Start of Problem and Control Information}
\label{start}

The first part of an input data file contains the {\it control data}
which consists of two records:
\begin{enumerate}
\item
A start/title record which must have as the
first four non-blank characters {\sl FEAP} (either upper or lower case letters
may be used with the remainder used as a problem title.
\item
The second recond contains
problem size information consisting of:
\begin{enumerate}
\item NUMNP - Number of nodal points;
\item NUMEL - Number of elements;
\item NUMMAT - Number of material property sets;
\item NDM - Space dimension of mesh;
\item NDF - Maximum number of unknowns per node; and
\item NEN - Maximum number of nodes per element.
\end{enumerate}
\end{enumerate}
As noted above,
input records for {\sl FEAP} are in free format.  Each data item
is separated by a comma, equal sign or blank characters.
If blank characters are used without commas, each data item {\it must} be
included.  That is multiple blank fields are not considered to be a zero.
Each data item is restricted to 14 characters (15 including the blank or
comma).

For standard input options
{\sl FEAP} can automatically determine the number of nodes,
elements, and material sets.  Thus, on the control record the values of NUMNP,
NUMEL, and NUMMAT may be omitted (i.e., specified as zero).  When using
automatic numbering it is generally advisable to use mesh input options
which avoid direct specification of a node or element number.  Specification
of nodal loads (forces), nodal displacements (displacements), and boundary
condition restraint codes have options which begin with {\tt E} and {\tt C}
for {\it edge} and {\it coordinate} related options, respectively.
It is recommended these be used whenever possible.

The use of the automatic determination of data requires a the mesh data
to be read twice: Once to do the counts and once to input the data.
For problems with a large number of data records,
this may result in significant time lapse during the input data phase.
The need for a second read may be avoided by inserting a {\tt NOCOunt}
record {\it before} the {\tt FEAP} record and providing the actual number of
nodes, elements and material sets on the control record.

We next consider commands used to describe the remainder of the finite
element mesh.  In {\sl FEAP} each data set starts with a command name
of which only the first four characters are used as identifiers.
Appendix A describes options for each mesh input and manipulation command.
Immediately following each command is the data to be processed.  Where 
a variable number of records is needed to define the data set a blank
line is used as a termination record.  Extra blank lines before or after
data sets are ignored.

Commands may be in any order.  If there is any order dependence {\sl FEAP}
will transfer the input data to temporary files and process it after the
mesh specification is terminated by the {\tt END} command.  Thus, information
will not necessarily occur in the output file in the order which data is
specified in the input file.

All data from a mesh input is written to the output file by default.  For very
large problems the size of the output file may become large.  Once a mesh
has been checked for correctness it may not be necessary to retain this
information in subsequent analyses.  Control of the data retained in the
output file is provided by using the {\tt PRIN}t and {\tt NOPR}int commands.
By default {\tt PRIN}t is assumed and all data is written.  Insertion of a
{\tt NOPR}int record before any data set (not within a data set) suspends
writing the data to the output file until another {\tt PRIN}t command is
encountered.
