\chapter[Material Models]{MATERIAL MODELS}
\label{matmods}

The data input for each of the current material options is summarized
below.  Tables are included to indicate which elements types can use
each type of data option.  As much as possible a common format and
notation is used for all the element types.

\section{Orthotropic Linear Elastic Models}
\label{orthel}

The orthotropic linear elastic material model in {\sl FEAPpv} is
given by
\begin{equation}
 \hat{\Bf{\epsilon}}
~=~ \hat{\Bf{C}} \, \hat{\Bf{\sigma}} ~+~ \hat{\Bf{\epsilon}}^{th}
\end{equation}
where $\hat{\Bf{\epsilon}}$ and $\hat{\Bf{\sigma}}$ are the stress and strain
arrays in the principal material directions and the elastic compliance array
in principal material directions is:
\begin{equation}
\hat{\Bf{C}} ~=~ \left[
\begin{matrix}
~\frac{1}{E_1} & - \frac{\nu_{12}}{E_1} & - \frac{\nu_{13}}{E_1} & 0 & 0 & 0 \\
- \frac{\nu_{21}}{E_2} & ~\frac{1}{E_2} & - \frac{\nu_{23}}{E_2} & 0 & 0 & 0 \\
- \frac{\nu_{31}}{E_3} & - \frac{\nu_{32}}{E_3} & ~\frac{1}{E_3} & 0 & 0 & 0 \\
0 & 0 & 0 & \frac{1}{G_{12}} & 0 & 0 \\
0 & 0 & 0 & 0 & \frac{1}{G_{23}} & 0 \\
0 & 0 &0 & 0 & 0 & \frac{1}{G_{31}} \\
\end{matrix} \right]
\end{equation}
with $E_i$ elastic moduli in principal directions, $\nu_{ij}$ Poisson ratios
for strains measured in the principal directions.  The above sign convention
corresponds to 
\begin{displaymath}
C_{ii} = \frac{1}{E_i} \quad \mbox{and} \quad
C_{ij} = - \, \frac{\nu_{ij}}{E_i} \quad \mbox{for} \quad i,j = 1,2,3
\end{displaymath}
and the definition of terms is identical to that given by
Christensen\scite{christensen71a} (except for shear modulus terms).

The thermal strain is given by:
\begin{equation}
\hat{\Bf{\epsilon}}^{th} ~=~ \left[
\begin{matrix}
\alpha_1 \\ \alpha_2 \\ \alpha_3 \\ 0 \\ 0 \\ 0 \\
\end{matrix} \right]
\, \Delta T ~=~ \hat{\Bf{\alpha}} \, \Delta T
\end{equation}
where
\begin{equation}
\Delta T ~=~ T\, - \, T_0 \, ,
\end{equation}
$\alpha_i$ are coefficients of linear thermal expansion and
$T_0$ is a specified reference temperature.

The orthotropic material parameters are input as shown in Table \ref{tab131a}
using the commands {\tt ELAStic,ORTHotropic} and {\tt THERmal,ORTHotropic}.
For 2-dimensional analyses the values of $G_{23}$ and $G_{31}$ are not used
and may be omitted.
The angle the principal directions makes with the $x_1$ (or $x$)
axis for plane stress
and plane strain analyses or the $r$ axis for axisymmetric analysis may be
specified using the material {\tt ANGLe} command
as shown in Table \ref{tab134a}.
Using this angle {\sl FEAPpv} transforms the input material compliances to
\begin{equation}
\Bf{C} = \Bf{R}^T \, \hat{\Bf{C}} \, \Bf{R}
\end{equation}
and converts the constitutive equation to the form
\begin{equation}
\Bf{\sigma} ~=~ \Bf{D} \, \Bf{\epsilon} ~+~ {\Bf{\beta}}^{th}
\end{equation}
where
\begin{equation}
\Bf{C} = \Bf{C}^{-1}
\end{equation}
and
\begin{equation}
{\Bf{\beta}}^{th} = - \, \Bf{D} \, {\Bf{\epsilon}}^{th}
\end{equation}

Material data is given by the command set:
\begin{verbatim}
       MATErial,1
         SOLId
         ELAStic ORTHotropic e1 e2 e3 nu12 nu23 nu31 g12 g23 g31
         THERmal ORTHotropic a1 a2 a3 t0
                                  ! blank termination record
\end{verbatim}
Additional data options and parameters are defined in Table \ref{tab131a}.

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | l | l |} \hline
Command & Type & Parameters \\ \hline
ELAStic & ORTHotropic & $E_1, E_2, E_3, \nu_{12},
\nu_{23}, \nu_{13}, G_{12}, G_{23}, G_{31}$ \\
ELAStic & ISOTropic & $E, \nu$ \\
ELAStic & TRANsverse & $E_1, E_2, \nu_{12}, \nu_{13}, G_{31}$ \\
DAMPing & RAYLeigh  & a0 , a1 \\
PLAStic & MISEs       & $Y_0, Y_{\infty}, \beta$ \\
PLAStic & HARDening   & $H_{iso}, H_{kin}$ \\
VISCoelastic & & $\mu_i, \tau_i$ \\
THERmal & ORTHotropic & $\alpha_1, \alpha_2, \alpha_3, T_0$ \\
THERmal & ISOTropic   & $\alpha, T_0$ \\ \hline
FOURier & ORTHotropic & $K_1, K_2, K_3, c$ \\
FOURier & ISOTropic   & $K, c$ \\ \hline
DENSity &       & $\rho$ \\ \hline
ANGLe   &       & $\psi$ \\ \hline
\end{tabular}
\end{center}
\caption{Material Model Data Inputs}
\label{tab131a}
\end{table}

\section{Isotropic Linear Elastic Models}
\label{isosel}

The isotropic models require less data since now only two independent
elastic parameters are needed to define $\hat{\Bf{C}}$. These are
taken as Young's modulus, $E$, and Poisson's ratio, $\nu$ and
for an isotropic material the elastic compliance array is
\begin{equation}
\hat{\Bf{C}} ~=~ \left[
\begin{matrix}
~\frac{1}{E} & - \frac{\nu}{E} & - \frac{\nu}{E} & 0 & 0 & 0 \\
- \frac{\nu}{E} & ~\frac{1}{E} & - \frac{\nu}{E} & 0 & 0 & 0 \\
- \frac{\nu}{E} & - \frac{\nu}{E} & ~\frac{1}{E} & 0 & 0 & 0 \\
0 & 0 & 0 & \frac{1}{G} & 0 & 0 \\
0 & 0 & 0 & 0 & \frac{1}{G} & 0 \\
0 & 0 &0 & 0 & 0 & \frac{1}{G} \\
\end{matrix} \right]
\end{equation}
with the shear modulus related through
\begin{equation}
G ~=~ \frac{E}{2 \, (1 \,+ \, \nu)}
\end{equation}
For thermally isotropic materials the expansion coefficient is constant in
all directions,thus
\begin{equation}
\hat{\Bf{\epsilon}}^{th} ~=~ \left[ \begin{matrix}
\alpha \\ \alpha \\ \alpha \\ 0 \\ 0 \\ 0 \\
\end{matrix} \right]
\, \Delta T
\end{equation}
The data input for the isotropic models is input using
the {ELASt\-ic,ISOTrop\-ic}
and {\tt THERm\-al,ISOTrop\-ic} commands as shown in Table \ref{tab131a}.
For an isotropic material it is not necessary to perform transformation of
the elastic arrays since $\Bf{C} ~=~ \hat{\Bf{C}}$.

The types of elements for which elastic material models may be specified
is indicated in Table \ref{tab131b}.

\section{Isotropic Finite Deformation Elastic Models}
\label{isofel}

Finite deformation hyperelastic models are provided in {\sl FEAPpv} for
several stored energy functions which are written in terms of deformation
measures.

Deformation measures may be defined in terms of positions in the reference
configuration, denoted by $\Bf{X}$, and positions in the current
configuration, denoted by $\Bf{x}$.  The motion of a point from the
reference to the current configuration at time $t$ is expressed as
\begin{equation}
\Bf{x} = \Bf{\varphi} ( \Bf{X}, t )
\end{equation}
The deformation gradient is then defined as
\begin{equation}
\Bf{F} = \frac{\partial \Bf{\varphi}}{\partial \Bf{X}}
\end{equation}

Additional measures of deformation are given by the right Cauchy-Green
deformation tensor
\begin{equation}
\Bf{C} = \Bf{F}^T \Bf{F}
\end{equation}
and the left Cauchy-Green deformation tensor
\begin{equation}
\Bf{b} = \Bf{F} \, \Bf{F}^T
\end{equation}
A measure of strain is provided by the Green strain
\begin{equation}
\Bf{E} = \frac{1}{2} \left( \Bf{C} - \Bf{1} \right)
\end{equation}

The hyperelastic model expressed in terms of the strain energy function
as a function of $\Bf{C}$ is given by
\begin{equation}
\Bf{S} = \frac{\partial W(\Bf{C})}{\partial \Bf{C}}
\end{equation}
where $W$ is a {\it stored energy} function.
Stress in the current configuration may be deduced by transformation (pushing)
the stress.  Accordingly
\begin{equation}
\Bf{\sigma} = \frac{1}{J} \, \Bf{F} \, \Bf{S} \, \Bf{F}^T
\end{equation}
Isotropic models may be
expressed in terms of the invariants of the deformation tensor.  Accordingly,
the three principal invariants given by
\begin{equation}
I_C = {\rm tr} \, \Bf{C}
\end{equation}
\begin{equation}
II_C = \frac{1}{2} \left( I_C^2 - {\rm tr} \, \Bf{C}^2 \right)
\end{equation}
and
\begin{equation}
III_C = \det \Bf{C} = J^2
\end{equation}
where $J$ is $\det \bf{F}$ may be used to write the stored energy function.

In {\sl FEAPpv} the isotropic elastic moduli are defined to match results from
the small strain isotropic elastic models. Accordingly, they only
require specification of
the elastic modulus, $E$, and Poisson ratio, $\nu$ or, equivalently, the
bulk modulus, $K$, and shear modulus, $G$.

\subsection{St. Venant-Kirchhoff}

The simplest model is a St. Venant-Kirchhoff model given by:
\begin{equation}
\Bf{S} = \Bf{D} \, \Bf{E}
\end{equation}
where $\Bf{S}$ is the second Piola-Kirchhoff stress, $\Bf{E}$ is the Green
strain, and $\Bf{D}$ are the elastic moduli.
This model may be deduced from the stored energy function
\begin{equation}
W = \frac{1}{2} \, \Bf{E}^T \Bf{D} \, \Bf{E}
\end{equation}
For isotropy the model may be written in terms of the invariants of $\Bf{E}$;
however, the $\Bf{D}$ will have the same structure as in an isotropic linear
elastic material (see above).

The St. Venant-Kirchhoff model should not be used
for problems
in which very large compressive deformations are expected.  The model gives
identical results to the small deformation isotropic model if deformations
are truly infinitesimal.  It is also a good model to use if the displacements
are large, but strains remain small.  For situations where large elastic
deformations are involved the {\tt NEOH}ookean or {\tt MNEO}hookean
models should be used.

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | l | l |} \hline
Command & Type & Parameters \\ \hline
ELAStic & NEOHookean & $E, \nu$ \\
ELAStic & MNEOhookean & $E, \nu$ \\
ELAStic & STVK & $E, \nu$ \\ \hline
\end{tabular}
\end{center}
\caption{Finite Deformation Elastic Material Data Inputs}
\label{tab131c}
\end{table}

\subsection{Neo-Hookean and Modified Neo-Hookean Models}

The stored energy functions for the finite deformation hyperelastic models
are split into two parts.  The first part defines the behavior associated
with volume changes and the second the behavior for other deformation states.
The volumetric deformation part is defined by a function $U(J)$ multiplied
by a material parameter.  The volumetric function is taken as
\begin{equation}
U(J) = \left( ln J \right)^2
\end{equation}

The neo-Hookean hyperelastic model is deduced from the stored energy function
\begin{equation}
W = \left( K  - \frac{2}{3} G \right) \, U(J)
+ \frac{1}{2} \, G \, \left( I_C - 3 \right)
\end{equation}
The parameters $K$ and $G$ are equivalent to the small strain bulk and shear
moduli, respectively.  Input data for the model is specified in terms of
the equivalent small strain modulus ($E$) and Poisson ratio ($\nu$) such that
the $K$ and $G$ are given by
\begin{equation}
\label{enu}
K = \frac{E}{3 \, ( 1 - 2 \nu )} ~~~~;~~~~ G = \frac{E}{2 \, ( 1 + \nu )}
\end{equation}
The data set to use this form is given by
\begin{verbatim}
       MATErial,1
         SOLId
         FINIte
         ELAStic NEOHook E nu
                                  ! blank termination record
\end{verbatim}

A modified form to the neo-Hookean model is also available.  The modified
form defines a stored energy function which splits the two terms into
pure volumetric behavior and pure deviatoric behavior.  To accomplish
the construction the deformation gradient is split into the product of
a volumetric and deviatoric form as
\begin{equation}
\Bf{F} = \Bf{F}_{vol} \, \Bf{F}_{dev}
\end{equation}
where
\begin{equation}
\Bf{F}_{vol} = J^\frac{1}{3} \Bf{1}
\end{equation}
and
\begin{equation}
\Bf{F}_{dev} =  {J}^{- \, \frac{1}{3}} \Bf{F}
\end{equation}
The stored energy function is then given as
\begin{equation}
W = K  \, U(J)
+ \frac{1}{2} \, G \, \left( {J}^{- \, \frac{2}{3}} I_C - 3 \right)
\end{equation}

The parameters $K$ and $G$ are again specified by their small strain
equivalent $E$ and $\nu$ defined in Eq. \ref{enu}.

The data set to use the modified form is given by
\begin{verbatim}
       MATErial,1
         SOLId
         FINIte
         ELAStic MNEOhook E nu
                                  ! blank termination record
\end{verbatim}

\section{Rayleigh Damping}
\label{raydamp}

The effects of damping may be included in transient solutions
assuming a damping matrix in the form
\begin{equation}
\Bf{C} ~=~ a_0 \, \Bf{M} ~+~ a_1 \, \Bf{K}
\end{equation}
This defines a form called {\it Rayleigh Damping}.
The input for this form of damping is given by:
\begin{verbatim}
       MATERIAL
         .....
           DAMPing RAYLeigh a0 a1
\end{verbatim}

This command is only included for small deformation elements using a
linear elastic material model and is used only for
time dependent solutions specified by a {\tt TRANsient} solution command.

\section{Viscoelastic Models}
\label{viscmod}

Materials which behave in a time dependent manner require extensions of the
elastic models cited above.  One model is given by viscoelasticity where
stress may be related to strain through either differential or integral
constitutive models (e.g., see {\sl FEAP} Theory Manual).\scite{feapt}  At present, the
implementation in {\sl FEAPpv} is restricted to isotropic viscoelasticity for
small deformation applications only in
which time effects are included for the deviatoric stress components only.
If we split the stress as:
\begin{equation}
\Bf{\sigma} ~=~ \sigma_{vol} \, \Bf{1} + \Bf{\sigma}_{dev}
\end{equation}
where $\sigma_{vol}$ represents the spherical part given by
$\frac{1}{3} \sigma_{kk}$ and $\Bf{\sigma}_{dev}$ is the deviatoric stress part.
Similarly the strain may be split as
\begin{equation}
\Bf{\epsilon} ~=~ \frac{1}{3} \, \theta \, \Bf{1} + \Bf{\epsilon}_{dev}
\end{equation}
where $\theta$ is the trace of the strain ($\epsilon_{kk}$)
and $\Bf{\epsilon}_{dev}$ is the deviatoric part.

The constitutive equation may now be written as
\begin{equation}
\Bf{\sigma}_{dev} ~=~ 2 \, G \, \int_{-\infty}^t
\mu(t - \tau ) \, \frac{d \Bf{\epsilon}_{dev}}{d \tau} \, d \tau
\end{equation}
where $mu(t)$ is a relaxation function.  The term $G mu(t)$ is called the
relaxation modulus.   In {\sl FEAPpv} the relaxation function is represented
by a Prony series
\begin{equation}
\mu(t) ~=~  \mu_0 + \sum_i \, \mu_i \, \exp \frac{-t}{\tau_i}
\end{equation}
The $\tau_i$ are time parameters defining the relaxation times
for the material and the $\mu_i$ are constant terms.  Currently, {\sl FEAPpv}
limits the representation to three (3) exponential terms.
The value of $\mu_0$ is computed from
\begin{equation}
\mu_0 ~=~  1 - \sum_i \, \mu_i
\end{equation}
Thus, the elastic modulus $G$ represents the instantaneous elastic response
and $G \, \mu_0$ the equilibrium, or long time, elastic modulus.
Only positive $\mu_i$ are permitted and care must
be taken in defining the $\mu_i$ to ensure that $\mu_0$ is positive or zero.
If $\mu_0$ is zero the response can have steady creep and never reach an
equilibrium configuration.

Input data for a one term model is given by the following data set:
\begin{verbatim}
       MATErial,1
         SOLId
         ELAStic ISOTropic 30e+06 0.3
         VISCoelastic term1 0.7   10.0
                                  ! blank termination record
\end{verbatim}
Here $\mu_1$ is 0.7 giving a $\mu_0$ of 0.3.  The relaxation time
is 10 time units.

After defining the response by the above exponential representation, the
constitutive equations are integrated in time by assuming the
strain rate is constant over each time step.  The method for integration
uses exact integration over each time step and leads to a simple recursion
for each exponential term (e.g., see \cite{taylorpg}).  Additional details
are also given in the {\sl FEAP} Theory manual.\scite{feapt}

\section{Plasticity Models}
\label{plasmod}

Classical elasto-plastic material models are included in {\sl FEAPpv} for
small deformation problems.
The stress for an elasto-plastic material may be computed by assuming
an additive split of the strain as
\begin{equation}
\Bf{\epsilon} = \Bf{\epsilon}^{el} + \Bf{\epsilon}^{pl}
\end{equation}

An associative flow rule is assumed so that the plastic strain rate
may be computed from a {\it yield function}, $F$, as
\begin{equation}
\dot{\Bf{\epsilon}}^{pl} = \dot{\gamma} \,
\frac{\partial F}{\partial \Bf{\sigma}}
\end{equation}
The relation may be integrated in time using a backward Euler (implicit)
time integration to compute a discrete form of the problem.

Isotropic and kinematic hardening are also added to the model.  The kinematic
hardening is limited to a linear form where it is assumed that
\begin{equation}
\Bf{\alpha} = H_{kin} \, \Bf{\epsilon}^{pl}
\end{equation}
where $\Bf{\alpha}$ is the back stress and $H_{kin}$ is the kinematic hardening
modulus.  The isotropic hardening is taken in a linear and saturation form
as
\begin{equation}
Y(e^{pl}) = Y_{\infty} + (Y_0 - Y_{\infty}) \exp (- \beta \, e^{pl})
+ H_{iso} \, e^{pl}
\end{equation}
where $Y_0$ is the initial uniaxial yield stress, $Y_{\infty}$ a stress
at large values of strain, $\beta$ a delay constant, and $H_{iso}$ is a linear
isotropic hardening modulus.  The accumulated plastic strain is computed
from
\begin{equation}
e^{pl} = \int_0^t \, \dot{\gamma} \, d \tau
\end{equation}

In {\sl FEAPpv} the discrete problem is solved using a closest point return
map algorithm (e.g., see \cite{simort85,simort86,simo:hughes}).

Input properties for a simple material with no saturation hardening and
linear isotropic hardening is given by:
\begin{verbatim}
       MATErial,1
         SOLId
         ELAStic ISOTropic 30e+06 0.3
         PLAStic MISEs     30e+03
         PLAStic HARDening 3000   0
                                  ! blank termination record
\end{verbatim}

\section{Heat Conduction Material Models}
\label{heat}

For thermal analysis a linear heat conduction capability is included in
{\sl FEAPpv}.  The constitutive equation is given by a linear Fourier
model in which the heat flux $\Bf{q}$ is related to the thermal gradient $\Bf{h}$
by the relation
\begin{equation}
\Bf{q} ~=~ - \, \Bf{K} \, \Bf{h}
\end{equation}
where, in the principal directions,
\begin{equation}
\Bf{K} ~=~ \left[ \begin{matrix}
K_1 & 0 & 0 \\
0 & K_2 & 0 \\
0 & 0 & K_3 \\
\end{matrix} \right]
\end{equation}
The values for $K_i$ and, for transient problems, the specific heat, $c$,
are specified using the command {\tt FOUR}\-ier,{\tt ORTH}o\-trop\-ic
or for the case where all are equal using {\tt FOUR}\-ier,{\tt ISOT}rop\-ic
as indicated in Table \ref{tab131a}

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c |} \hline
Command & Solid & Truss & Frame & Plate & Shell & Membrane & Thermal \\ \hline
ELAStic       & X & X & X & X & X & X & - \\
PLAStic       & S & X & - & - & - & - & - \\
VISCoelastic  & S & X & - & - & - & - & - \\
THERmal       & S & - & - & X & - & X & - \\
FOURier       & S & - & - & - & - & - & X \\
DENSity       & X & X & X & X & X & X & X \\
ANGLe         & X & - & - & X & X & X & X \\ \hline
\end{tabular}
\end{center}
\caption{Material Commands vs. Element Types. X=all, F=finite, S=small.}
\label{tab131b}
\end{table}

\section{Mass Matrix Type Specification}
\label{masstype}

The mass matrix for continuum problems and the specific heat matrix
for thermal problems may be either a {\it consistent}, {\it lumped},
or {\it interpolated} form.  By default {\it FEAPpv} uses a lumped matrix.
If $\Bf{M}_{cons}$ is the consistent matrix and $\Bf{M}_{lump}$ is the diagonal
lumped matrix, the interpolated matrix is defined as:
\begin{equation}
\Bf{M}_{interp} ~=~ (1 \,-\, a) \, \Bf{M}_{cons} ~+~ a \, \Bf{M}_{lump}
\end{equation}
The type of mass and, where required,
the parameter $a$ are input using the {\tt MASS} command as shown in
Table \ref{tab132a} and the elements which are affected by the command
are indicated in Table \ref{tab132b}.

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | l | l |} \hline
Command & Type & Parameters \\ \hline
MASS & LUMPed       & {}  \\
MASS & CONSistent   & {}  \\
MASS & {}           & $a$ \\ \hline
\end{tabular}
\end{center}
\caption{Material Model Mass Related Inputs}
\label{tab132a}
\end{table}

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c |} \hline
Command & Solid & Truss & Frame & Plate & Shell & Membrane & Thermal \\ \hline
MASS          & X & X & X & X & - & - & X \\ \hline
\end{tabular}
\end{center}
\caption{Mass Command vs. Element Types}
\label{tab132b}
\end{table}

\section{Element Cross Section and Load Specification}
\label{xsect}

\subsection{Resultant formulations}

The plane stress and structural elements require specification of
cross-section information.
For the plane stress, plate, and shell
elements this is a thickness which is specified using the 
\textbf{THICkness} command as shown in Table \ref{tab133a}.
The plate element also permits the effects
of transverse shear deformation to be included and, if this is
different than the $\kappa = 5/6$ default value,
it is also given using the thickness command.

For the truss element it is necessary to provide the cross-sectional area, $A$.
This is specified using the \textbf{CROSs section} command as indicated in
Table \ref{tab133a}.

For frame elements it is also necessary to provide cross-sectional
area and appropriate flexural effects as indicated in Table \ref{tab133a}.
For two dimensional frame elements only the area, $A$, flexural stiffness,
$I_{xx}$ and, for formulations which include shear deformation, $\kappa_x$
need be specified (if $\kappa_x$ is zero a default values of 5/6 is used).
Similarly, for the three-dimensional frame element no
shear deformation is included in the current implementation and, thus, no
values for $\kappa_x$ and $\kappa_y$ need be specified.\footnote{The output file
may include values for some unused parameters.}

Element loads for surface pressure and body force are input using the
{\tt LOAD} and {\tt BODY} force commands as shown in Table \ref{tab133a}.

The types of elements affected by the {\tt THICkness}, {\tt LOAD} and
{\tt BODY} commands is indicated in Table \ref{tab133b}.

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | l | l |} \hline
Command & Type & Parameters \\ \hline
THICkness & {}      & $h, \kappa$ \\
CROSS     & section & $A, I_{xx}, I_{yy}, I_{xy},
J_{zz}, \kappa_x, \kappa_y$ \\ \hline
BODY & forces   & $b_1, b_2, b_3$ \\
LOAD & normal   & $q$ \\ \hline
\end{tabular}
\end{center}
\caption{Cross Section and Body Force Inputs}
\label{tab133a}
\end{table}

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c |} \hline
Command & Solid & Truss & Frame & Plate & Shell & Membrane & Thermal \\ \hline
THICkness     & X & - & - & X & X & X & X \\
CROSs         & - & X & X & - & - & - & - \\ \hline
BODY          & X & X & X & - & X & X & - \\
LOAD          & - & - & - & X & X & X & - \\ \hline
\end{tabular}
\end{center}
\caption{Geometry and Loads vs. Element Types}
\label{tab133b}
\end{table}

\section{Miscellaneous Material Set Parameter \\ Specifications}
\label{miscmat}

In addition to the above material, geometric and loading parameters the
values for some other variables may also be set.

It is possible to replace global parameters for the type of two dimensional
analysis using the {\tt PLANe STREss}, {\tt PLANe STRAin}, or {\tt AXISymmetric}
commands.  Similarly the global value for the temperature degree of
freedom to use in coupled thermo-mechanical problems may be changed
for the current material set using the {\tt TEMPerature} command.  The
formats are indicated in Table \ref{tab134a} and the affected element types in
Table \ref{tab134b}.
The values for the number of quadrature points (in elements, not cross sections)
to be used for computing arrays and element
outputs may be set using the {\tt QUADrature} command.
Generally, {\sl FEAPpv} will select an appropriate order of quadrature to
be used in computing the arrays and for output of element quantities.  Thus,
care should be used in changing the default values.

{\sl FEAPpv} includes capabilities to solve finite deformation problems
using only the 2-d {\tt SOLId}, {\tt FRAMe}, and {\tt TRUSs}
elements.  To select the
finite deformation element it is necessary to use the {\tt FINIte} deformation
option instead of the default {\tt SMALl} deformation option.
This may be done for all materials using the {\tt GLOBal} command.
There are three different element technologies
which may be selected {\tt DISPlacement} (which is the default) or {\tt MIXEd}
types. The data options for these are indicated
in Table \ref{tab134a} and the affected element types in Table \ref{tab134b}.

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | l | l |} \hline
Command & Type & Parameters \\ \hline
QUADrature  & {}       & $n_{array}, n_{output}$ \\
PENAlty     &          & $k_{pen}$ \\
TEMPerature & {}       & $T_{dof}$ \\ \hline
SMALl       & deformation & {} \\
FINIte      & deformation & {} \\
NONLinear   & {}       & {} \\ \hline
DISPlacment & {}       & {} \\
MIXEd       & {}       & {} \\
PLANe       & STREss   & {} \\
PLANe       & STRAin   & {} \\
AXISymmetric & {}      & {} \\ \hline
\end{tabular}
\end{center}
\caption{Miscellaneous Material Model Inputs}
\label{tab134a}
\end{table}

\begin{table}[ht!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c |} \hline
Command & Solid & Truss & Frame & Plate & Shell & Membrane & Thermal \\ \hline
QUADrature      & X & - & -  & - & X & X & X \\
PENAlty         & - & - & -  & - & - & - & - \\
TEMPerature     & X & X & X  & X & X & - & - \\ \hline
SMALl           & X & X & X  & - & X & X & - \\
FINIte          & X & X & X  & - & X & X & - \\
NONLinear       & - & X & X  & - & - & - & - \\ \hline
DISPlacement    & X & - & -  & - & - & - & - \\
MIXEd           & X & - & -  & - & - & - & - \\
ENHAnced        & X & - & -  & - & - & - & - \\ \hline
PLANe STREss    & X & - & -  & - & - & - & X \\
PLANe STRAin    & X & - & -  & - & - & - & X \\
AXISymmetric    & X & - & -  & - & - & - & X \\ \hline
\end{tabular}
\end{center}
\caption{Miscellaneous Material Commands vs. Element Types}
\label{tab134b}
\end{table}
